INGRID: Creating Languages in MPS from ANTLR Grammars

### 2017-06-07 chapters/background.tex
new format
notes about projectional editors

@@

\begin{figure}[ht]
\centering
\begin{alltt}
\small
\mpstgkeyword{text gen component for concept} \mpstgtarget{IfStatement} \{
  \mpstgparam{(context, buffer, node)->void} \{
    \mpstgaction{append} \textcolor{Blue}{\textbf{\textbackslash{}n;}}
    \mpstgaction{indent buffer;}
    \mpstgaction{append} \{\mpstgliteral{if (}\} \$\{\mpstgparam{node}.\mpstgnodeprop{condition}\} \{\mpstgliteral{) \{}\};
    \mpstgkeyword{with indent} \{
      \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{ifTrue}\};
    \}
    \mpstgaction{append} \textcolor{Blue}{\textbf{\textbackslash{}n}} \{\mpstgliteral{\}}\} \$list\{\mpstgparam{node}.\mpstgnodeprop{elsifClauses}\};
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{ifFalseStatement}.\textbf{isNotNull}) \{
      \mpstgaction{append} \{ \mpstgliteral{else}\} \$\{\mpstgparam{node}.\mpstgnodeprop{isFalseStatement}\};
    \}
  \}
\}
\end{alltt}
\caption{TextGen aspect definition for the \code{if-else} statement}
\label{fig:if_statement_textgen}
\end{figure}

>>>>

\todo{Napsat sem nebo pripadne do Introduction (ale radsi sem): MPS platform, and tools built upon it (e.g., mbeddr) addresses to a large degree known limitations of projectional editors, such as usability of editors for languages based on textual notation (e.g., non-linear typing) and support for version control [citovat "M. Voelter, J. Siegmund, T. Berger, and B. Kolb. Towards User-Friendly Projectional Editors. SLE 2014, LNCS 8706"].}

\todo{Mozna pouzit termin "language concept" misto samotneho "concept".}

@@

\begin{figure}[ht]
\centering
\begin{alltt}
\small
\mpstgkeyword{text gen component for concept} \mpstgtarget{IfStatement} \{
  \mpstgparam{(context, buffer, node)->void} \{
    \mpstgaction{append} \textcolor{blue}{\textbf{\textbackslash{}n;}}
    \mpstgaction{indent buffer;}
    \mpstgaction{append} \{\mpstgliteral{if (}\} \$\{\mpstgparam{node}.\mpstgnodeprop{condition}\} \{\mpstgliteral{) \{}\};
    \mpstgkeyword{with indent} \{
      \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{ifTrue}\};
    \}
    \mpstgaction{append} \textcolor{blue}{\textbf{\textbackslash{}n}} \{\mpstgliteral{\}}\} \$list\{\mpstgparam{node}.\mpstgnodeprop{elsifClauses}\};
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{ifFalseStatement}.\textbf{isNotNull}) \{
      \mpstgaction{append} \{ \mpstgliteral{else}\} \$\{\mpstgparam{node}.\mpstgnodeprop{isFalseStatement}\};
    \}
  \}
\}
\end{alltt}
\caption{TextGen aspect definition for the \code{if-else} statement}
\label{fig:if_statement_textgen}
\end{figure}

### 2017-06-07 chapters/introduction.tex
notes about projectional editors

Projectional editing, keeping the code in the AST form, and the absence of parsers, brings along several benefits.
\begin{itemize}
	\item The developers practically cannot enter syntactically invalid code, because a projectional editor controls the interaction between the user and the code.
		%An AST created by the editor can represent only a syntactically valid code.
	\item Programming languages can be defined in a modular way, and multiple languages can be easily combined together or one can extend another, while parsers seriously limit language modularity for traditional parsed languages.
	\item The languages may support contextual or non-parseable notations, such as tables, diagrams and positional syntax.
	\item Since the projection is detached from the physical representation of code (AST), authors of languages can define multiple notations and allow the developers to switch between different representations of the code on the screen.
\end{itemize}
All of this is useful especially for DSLs, which are frequently used by domain experts who may not be professional software developers.
The mbeddr project~\cite{ref:MBEDDR}, which extends C with numerous domain-specific constructs and data types, illustrates the abilities of language workbenches in general and of MPS in particular.

>>>>

Projectional editing, keeping the code in the AST form, and the absence of parsers, brings along several benefits.
\begin{itemize}
	\item The developers practically cannot enter syntactically invalid code, because a projectional editor controls the interaction between the user and the code.
		%An AST created by the editor can represent only a syntactically valid code.
	\item Programming languages can be defined in a modular way, and multiple languages can be easily combined together or one can extend another, while parsers seriously limit language modularity for traditional parsed languages.
	\item The languages may support contextual or non-parseable notations, such as tables, diagrams and positional syntax.
	\item Since the projection is detached from the physical representation of code (AST), authors of languages can define multiple notations and allow the developers to switch between different representations of the code on the screen.
\end{itemize}
All of this is useful especially for DSLs, which are frequently used by domain experts who may not be professional software developers.
The mbeddr project~\cite{ref:MBEDDR}, which extends C with numerous domain-specific constructs and data types, illustrates the abilities of language workbenches in general and of MPS in particular.
\todo{Popsat dalsi benefits of projectional editors/IDEs (and of MPS in particular): Projectional editors never have to parse the concrete syntax, so they can support/use more complicated syntax/layout. Support for non-linear notations (mathematical expressions, graphical or tabular layout), notations that use two-dimensional layout. Projectional editors mix different notations/styles.}
\todo{Nekam pridat text: Benefits and limitations of projectional editors are thoroughly discussed in [M. Voelter, J. Siegmund, T. Berger, and B. Kolb. Towards User-Friendly Projectional Editors. SLE 2014, LNCS 8706].}
